---
title: "Compile Filtered Dataset - R Notebook"
authors:  Rebecca Ross and Genoveva Gonzalez Mirelis
date: "Last Rendered on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 2
    toc_float: yes
    fig_width: 7
    fig_height: 7
always_allow_html: true 
---

#Assumptions
This script allows you to run species/ station filters/ any other filters in what ever order suits you, then compile the final objects here to ensure you are getting the up-to-date filters applied.

Assumes:
- pre run config
- pre run Filters scripts


#Libraries
```{r}
library(dplyr)
```


#Combine filters to form dataset

Use output of stationFilters (sample_info_f_ss) to select from filtered species data (sel.sppDens)
```{r}
stFilt.sel.sppDens<-sel.sppDens%>% filter(SampID%in%sample_info_f_ss$SampID) 

length(unique(sel.sppDens$SampID)) # number of samples in sp filtered dataset
length(unique(stFilt.sel.sppDens$SampID)) # number of samples in station filtered sp filtered dataset
```

#Change to wide format for analyses
```{r}

```

