---
title: "Data quality filters, and subsetting the data for manageability"
authors: Genoveva Gonzalez Mirelis and Rebecca Ross
date: "Last Rendered on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 2
    toc_float: yes
    fig_width: 7
    fig_height: 7
always_allow_html: true 
---

# Introduction

In this notebook we filter out stations and/or samples that fail various data quality tests. We also subset the data to make it more or less one third of the total.

Requirements: "sppdens", "sample_info", "refer", "MAREANO_GBK-sampled_areas"

Outputs: "sppdens_qc"

# Libraries and input files

```{r}
library(rgdal)
library(tidyverse)
#library(ggplot2)

# read needed shapefile
dsn <- "U:/Mareano/VIDEOLAB/VIDEO DATA/200m_scale_species_by_sample/GIS_Data"
gbk <- readOGR(dsn = dsn, layer = "MAREANO_GBK-sampled_areas_311221") 
  #%>% fortify

```

# Station-level quality assessment

## Total number of Ref stations
This includes all stations completed in the field up to cruise number below
```{r}
dim(refer)[1]
```
#### Last cruise included in this lot is...
```{r}
max(refer$cruise_no)
```

## List stations whose data is not in MarVid
Check for reasons why. Keep in mind that it's possible the station is yet to be analyzed.
```{r}
x<-refer %>% filter(is.na(total_records))%>%
  select(nRefStation,nSampleNumber,Comments,Other.comments,Notes,Notes.II, lon_mid_dec, lat_mid_dec)
x
```
## View the spatial distribution of stations not in MarVid
```{r}
# convert reference to spatial points
x_spat <- x %>%
  select(lon_mid_dec,lat_mid_dec) %>%
  SpatialPoints %>%
  SpatialPointsDataFrame(x) %>%
  utmize

par(mar = c(0, 0, 0, 0) + 0.1)
plot(gbk, border="green")
plot(x_spat, axes = FALSE, cex = 0.6, add = TRUE) 
box() 
```

## Assess data completeness
We should use only video lines where data is complete. One way to determine that is by looking at the total number of biological annotations relative to the towed distance. Note that total number of biological annotations depends on species richness, so expect large variation!
```{r}
hist(refer$total_bio/refer$towed_distance, main = "Distribution of records:distance ratio",
     xlab = "total records : towed distance")
```

### Stations with an unexpected number of annotations in relation to length
Check how much data we would be using/discarding for each threshold
```{r}
thres <- 5
good <- refer$nSampleNumber[which(refer$total_bio/refer$towed_distance<thres)]
not_good <- refer %>% 
  filter(!(nSampleNumber%in%x$nSampleNumber)) %>%
  filter(!(nSampleNumber%in%good))
not_good %>%
  select(nRefStation,nSampleNumber,towed_distance,total_records)
```
##### I have checked these and they are either fine, or will fall out later on. Turn this one off

## Filter stations for other reasons
```{r}
refer %>%
  filter(!is.na(reason_excluded)) %>%
  select(nRefStation,nSampleNumber,reason_excluded)
```

## Final set of Ref Stations
Count
```{r}
stationFilter <- refer %>%
  filter(!(nSampleNumber%in%x$nSampleNumber)) %>% # eliminate those whose data is not in MarVid
  #filter(nSampleNumber%in%good) %>%               # eliminate those with weird no. of bio annotations
  filter(is.na(reason_excluded)) %>%              # mixed reasons to be excluded (short lines, wiggly lines, etc.)
  select(nSampleNumber) %>%
  pull()
length(stationFilter)
```

# Sample-level quality assessment

## Count of Video Lines that have been split into Samples
Only navigation data is required for this
```{r}
length(unique(sample_info$VL))
```
### Video Lines left after applying stationFilter
```{r}
sample_info %>%
  select(VL) %>%
  unique() %>%
  filter(VL%in%stationFilter) %>%
  pull %>%
  length
```

## Examine distribution of surveyed distance and area
Examine the distribution of total area and length of samples
```{r}
sample_info_f <- sample_info %>% 
  filter(VL%in%stationFilter) %>%
  mutate(area_m2 = SegLengthM * meanFoV_m)
par(mfrow=c(1,2))
hist(sample_info_f$area_m2, main ="Surveyed area", xlab = "Area (m2)")
hist(sample_info_f$SegLengthM, main ="Surveyed distance", xlab = "Length (m)")
```

## Set thresholds for data quality
Decide what is acceptable in terms of length of sample and picture quality. The percentages shown below are relative to the whole set of MAREANO stations (and not the "good" stations!)
```{r, warning=FALSE}
# thresholds for total surveyed area
#survarea_upper <- 1200
#survarea_lower <- 250

# alternative, set thresholds for survey length (maybe better)
length_lower <- 200 
length_upper <- 250 

# threshold for picture quality
# in English, Doc: "either unknown, or not more than half with no vision, or not more than three quarters with poor vision"
poorvis_thres <- 0.5 
novis_thres <- 0.5

#table(ifelse(is.na(sample_info_f$ok_vision),1,
#                       ifelse(sample_info_f$poor_vision<poorvis_thres&sample_info_f$no_vision<novis_thres,1,0)))

filter1 <- sum(with(sample_info_f, SegLengthM>length_lower&SegLengthM<length_upper))
filter2 <- length(which(with(sample_info_f, ifelse(is.na(ok_vision),1,
                    ifelse(poor_vision<poorvis_thres&no_vision<novis_thres,1,0)))==1))
filter <- #sum(
    rowSums(
  cbind((with(sample_info_f, SegLengthM>length_lower&SegLengthM<length_upper)),ifelse(is.na(sample_info_f$ok_vision),1,
       with(sample_info_f, ifelse(poor_vision<poorvis_thres & no_vision<novis_thres,1,0)))))==2
#)

summary<-data.frame("length_filter_pass"=filter1/dim(sample_info_f)[1],
                    "picqual_filter_pass" =filter2/dim(sample_info_f)[1],
                    "both" = sum(filter)/dim(sample_info_f)[1])

summary
```

### Check out those that get filtered out
#### Length filter fail
```{r}
lengFilter <- with(sample_info_f,SegLengthM>length_lower&SegLengthM<length_upper)
sample_info_f %>% filter(!lengFilter)
```


```{r}
# convert reference to spatial points
sif_spat <- sample_info_f %>% filter(!lengFilter) %>%
  select(x_coord,y_coord) %>%
  SpatialPoints

par(mar = c(0, 0, 0, 0) + 0.1)
plot(gbk, border="green")
plot(sif_spat, axes = FALSE, cex = 0.6, add = TRUE) 
box() 
```

###### Spatial distribution

#### Picture quality filter fail
```{r}
picqFilter <-with(sample_info_f, ifelse(is.na(ok_vision),1,
                    ifelse(poor_vision<poorvis_thres&no_vision<novis_thres,1,0)))==1
sample_info_f %>% filter(!picqFilter)
```

```{r}
# convert reference to spatial points
sif_spat <- sample_info_f %>% filter(!picqFilter) %>%
  select(x_coord,y_coord) %>%
  SpatialPoints

par(mar = c(0, 0, 0, 0) + 0.1)
plot(gbk, border="green")
plot(sif_spat, axes = FALSE, cex = 0.6, add = TRUE) 
box() 
```
###### Spatial distribution

## Final set of samples
```{r}
combinedFilter <- filter

sample_info_f %>% 
  dplyr::filter(combinedFilter) %>%
  select(SampID) %>%
  pull %>%
  length


```
## See final distribution

```{r}
fin_spat <- sample_info_f %>% dplyr::filter(combinedFilter) %>%
  select(x_coord,y_coord) %>%
  SpatialPoints

par(mar = c(0, 0, 0, 0) + 0.1)
plot(gbk, border="green")
plot(fin_spat, axes = FALSE, cex = 0.6, add = TRUE) 
box() 
```

# Subset samples for maneagbility
## Take a random (sub)sample from each Ref station

```{r}
sample_info_f_ss <- sample_info_f %>% dplyr::filter(combinedFilter)  %>% group_by(VL) %>% slice_sample(n=1)
length(unique(sample_info_f_ss$VL))

```

write the subsample to csv with sys.Date appended to keep track of which is the active sample that day
```{r}
## export subsampled dataset with 1 sample per R station
write.csv(sample_info_f_ss, file=file.path(dataPath,(paste0("activeSubsample_", Sys.Date(),".csv"))))

## export whole dataset (i.e., without subsetting for manageability)
write.csv(sample_info_f, file=file.path(dataPath,(paste0("activeSubsampleOslo_", Sys.Date(),".csv"))))

```

